---
title: "Life_expectancy"
output: pdf_document
---

```{r message=FALSE}
library(rvest)
library(ggplot2)
library(dplyr)
library(maps)
library(scales)
```

## Import data from wikipedia
```{r}
life_expect <- read_html("https://en.wikipedia.org/wiki/List_of_U.S._states_by_life_expectancy")
```

## Convert data into a table
```{r}
life_expect <- life_expect %>%
  html_nodes(xpath = "//table") %>%
  .[[3]] %>%
  html_table(fill = T)
```

## Clean life_expect data frame by selecting required columns
```{r}
life_expect <- life_expect[, c(3:6)]
```

## Rename the columns
```{r}
names(life_expect)[c(1, 2, 3, 4)] <- c("region", "life_expect_all", "life_expect_White", "life_expect_Black")
```

## Convert black and white expectancy to numeric variables:
```{r warning=FALSE}
life_expect <- life_expect %>%
  mutate(
    life_expect_White = as.numeric(life_expect_White),
    life_expect_Black = as.numeric(life_expect_Black),
    region = tolower(region)
        )
```

## Compute the difference between black and white life expect:
```{r}
life_expect <- life_expect %>%
  mutate(life_expect_diff = life_expect_White - life_expect_Black)
```

## How many regions(states) where African Americans live more than white Americans:
```{r}
life_expect %>%
  filter(life_expect_diff < 0) %>%
  arrange(desc(life_expect_diff))
```
  There are 9 states where life_expect is larger for African Americans with the highest value for New Hamsphire.

## How many regions(states) where life expectancy is larger for white Americans than African Americans:
```{r}
life_expect %>%
  filter(life_expect_diff > 0) %>%
  arrange(desc(life_expect_diff))
```
  There are 42 states where life_expect is larger for white American with the highest value for District of Columbia.

## Load the map data:
```{r warning=FALSE}
states = map_data("state")
str(states)
```

## Merge the two datasets:
```{r warning=FALSE}
states <- merge(states, life_expect, by = "region", all.x = T)
str(states)
```

## Map of difference in life expectancy between White Americans & African Americans
```{r}
ggplot(data = states, aes(x = long, y = lat, group = group, fill = life_expect_diff)) + geom_polygon(col = "white") + 
  scale_fill_gradient(name = "Years", low = "#132B43", high = "#56B1F7", guide = "colorbar", na.value = "lightblue", breaks = pretty_breaks(n = 5)) + 
  labs(title = "Difference in Life Expectancy between African Americans & white Americans")
```

## Map of African Americans life expectancy
```{r}
ggplot(data = states, aes(x = long, y = lat, group = group, fill = life_expect_Black)) + geom_polygon(col = "white") + 
  scale_fill_gradient(name = "Years", low = "#ffe8ee", high = "#c81f49", guide = "colorbar", na.value = "lightblue", breaks = pretty_breaks(n = 5)) + 
  labs(title = "Life Expectancy for African Americans")
```

## Map of White Americans life expectancy
```{r}
ggplot(data = states, aes(x = long, y = lat, group = group, fill = life_expect_White)) + geom_polygon(col = "white") + 
  scale_fill_gradient(name = "Years", low = "#ffe8ee", high = "#c81f49", guide = "colorbar", na.value = "lightblue", breaks = pretty_breaks(n = 5)) + 
  labs(title = "Life Expectancy for White Americans")
```

 

## Map of US states by life expectancy 
```{r} 
ggplot(data = states, aes(x = long, y = lat, group = group, fill = life_expect_all)) + geom_polygon(col = "white") + 
    scale_fill_gradient(name = "Years", low = "#132B43", high = "#56B1F7", guide = "colorbar", na.value = "lightblue", breaks = pretty_breaks(n = 5)) + 
    labs(title = "Life Expectancy in the USA")
``` 

  